An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-05-08 EconGrowth.Rnw
Minor edits.

\noindent where the value of $Q_{m,workers}$ for the U.S. ($\sim$ 4200 kcal/day-worker) was taken from a time series of kcal/day-person for workers for 1980--2011 \citep{Warr:2012a}. $Q_{m,workers}$ for other countries was taken from the \citet{FAO:2013b}. The value for $N_{workers}$ was obtained from the labor time series for each economy as discussed in Section \ref{sec:Labor_Data}.

@@

\noindent $\eta_{food \rightarrow work, animals}$ and $\eta_{food \rightarrow work, workers}$ are animal and worker efficiency of converting food into energy respectively. The animal feed-to-work conversion efficiency ($\eta_{food \rightarrow work, animals}$ = 0.4) was taken from \citet{Warr:2012a} and assumed equal across all economies. The food-to-work efficiency for workers ($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from \citet{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug}, an additional factor ($f_{enduse/intake}$) was included to account for workers' food energy that sustains the body's normal function as opposed to providing energy for physical labor \citep{Wirsenius:2000vi}. $f_{enduse/intake}$ varies by country (Table \ref{tab:f_enduse_intake}). A similar enduse-to-intake ratio could also be applied to animal work, but it was unavailable from \citet{Wirsenius:2000vi}.\footnote{The error in exergy caused by neglecting $f_{enduse/intake}$ for animals is very small, because animal muscle exergy is a small fraction of the total energy supply to any economy, with a maximum of 4.1\% for China in 1992.}

\begin{table} \caption{Food enduse-to-intake ratios for all countries.} \label{tab:f_enduse_intake}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{enduse/intake}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}

@@

The conversion of a raw energy carrier (e.g., coal) into productive energy for an economy (e.g., electricity) necessarily involves inefficiencies. In the case of coal consumption, only about one-third of the coal exergy is converted to useful electricity: the rest is waste heat. The concept of ``useful work,'' pioneered by Ayres and Warr \citep{Ayres:2003ec, Ayres:2005kn}, build upon exergy analysis to consider efficiencies of both energy production and end use. To calculate useful work ($U$), the total exergy input to an economy ($X$) is decomposed into several end uses: high temperature industrial heat (e.g., steel production), medium temperature heat (e.g., steam production), low temperature space heat (for residential heating), mechanical drive (energy used for prime mover engines of all types except electric motors), electricity (excluding lighting), lighting (from both electricity and fuels), and muscle work (from animals and workers) as shown in the following equation:

@@

\noindent where the subscript $i$ indicates the energy end use, $\varepsilon_i$ is the efficiency of converting exergy into useful energy for each end use, and $\phi_i$ gives the fraction of exergy input to an economy that goes toward the i$^\mathrm{th}$ end use.

>>>>

\noindent where the value of $Q_{m,workers}$ for the U.S. ($\sim$ 4200 kcal/day-worker) was taken from a time series of kcal/day-person for workers for 1980--2011 \citep{Warr:2012a}. $Q_{m,workers}$ for other countries was taken from the \citet[``\texttt{Dietary Energy Cons.~Countries}'' tab]{FAO:2013b}. The value for $N_{workers}$ was obtained from the labor time series for each economy as discussed in Section \ref{sec:Labor_Data}.

@@

\noindent $\eta_{food \rightarrow work, animals}$ and $\eta_{food \rightarrow work, workers}$ are animal and worker efficiency of converting food into energy, respectively. The animal feed-to-work conversion efficiency ($\eta_{feed \rightarrow work, animals}$ = 0.4) was taken from \citet{Warr:2012a} and assumed equal across all economies. The food-to-work efficiency for workers ($\eta_{food \rightarrow work, workers}$ = 0.2) was taken from \citet[Table~B.1]{Ayres:2010ug}. Following \citet[p.~315]{Ayres:2010ug}, an additional factor ($f_{enduse/intake}$) was included to account for workers' food energy that sustains the body's normal function as opposed to providing energy for physical labor \citep{Wirsenius:2000vi} **** Caleb needs to find a detailed source (table, equation, number, page number) in Wirsenius for this approach *****. $f_{enduse/intake}$ varies by country (Table \ref{tab:f_enduse_intake}). A similar enduse-to-intake ratio could also be applied to animal work, but it was unavailable from \citet{Wirsenius:2000vi}.\footnote{The error in exergy caused by neglecting $f_{enduse/intake}$ for animals is very small, because animal muscle exergy is a small fraction of the total energy supply to any economy, with a maximum of 4.1\% for China in 1992.}

\begin{table} \caption{Food enduse-to-intake ratios ($f_{enduse/intake}$) for all countries.} \label{tab:f_enduse_intake}
  \begin{center}
    \begin{tabular}{r c c c c c c c c c}
      Country & US & UK & JP & CN & ZA & SA & IR & TZ & ZM \\
      \hline
      $f_{enduse/intake}$ & 0.62 & 0.65 & 0.76 & 0.76 & 0.87 & 0.72 & 0.72 & 0.72 & 0.72\\
    \end{tabular}
  \end{center}
\end{table}

@@

The conversion of a raw energy carrier (e.g., coal) into productive energy for an economy (e.g., electricity) necessarily involves inefficiencies. In the case of coal consumption, only about one-third of the coal exergy is converted to useful electricity: the rest is waste heat. The concept of ``useful work,'' pioneered by Ayres and Warr \citep{Ayres:2003ec, Ayres:2005kn}, builds upon exergy analysis to consider efficiencies of both energy production and enduse. To calculate useful work ($U$), the total exergy input to an economy ($X$) is decomposed into several end uses: high temperature industrial heat (e.g., steel production), medium temperature heat (e.g., steam production), low temperature space heat (for residential heating), mechanical drive (energy used for prime mover engines of all types except electric motors), electricity (excluding lighting), lighting (from both electricity and fuels), and muscle work (from animals and workers) as shown in the following equation:

@@

\noindent where the subscript $i$ indicates the energy enduse, $\varepsilon_i$ is the efficiency of converting exergy into useful energy for each enduse, and $\phi_i$ gives the fraction of exergy input to an economy that goes toward the i$^\mathrm{th}$ enduse.
